% $Id$

\section{Available hardware and software}
\label{sec:availabe-hardware-and-software}

This section lists the hardware I have used during development of
Cactus.  


\subsection{Asserballe}
\label{sec:asserballe}

Urls in long banes for this.

\begin{itemize}
\item Pentium-S 133 MHz PC with 128 Mb RAM and 10 Gb IDE disk
\item Linux Redhat 6.1  
\item \myurl{http://www.mysql.org}{MySQL} 3.22.22
\item \texttt{Apache/1.3.11 (Unix) PHP/3.0.15 ApacheJServ/1.1 mod\_perl/1.21
}
\item \myurl{http://xml.apache.org}{Cocoon 1.7.1} for rendering XML to
  HTML on the fly.  Can do SQL-requests, and render them easily to
  HTML.
\item \myurl{http://www.ibm.com/java/jdk/118/linux/}{IBM JDK 1.1.8 for
    Linux}.  This is currently the fastest and most reliable Java
  engine for Linux.
\item \myurl{http://www.php.net}{PHP} for dynamic web-pages with none
  or some SQL-data.
\item \myurl{http://www.perl.com}{Perl 5.6.0} for doing complex
  background tasks with DBI::MySQL perl modules for accessing
  the MySQL databases.
\end{itemize}

The choice of JDK was made at a time when
\myurl{http://developer.java.sun.com/developer/bugParade/top25bugs.html}{
  top bug at Javasoft Developer Connection} was ``please provide a
port of Java to Linux'', and where IBM beat Sun to it by porting their
Java 1.1.6 implementation to Linux.  The benchmarks showed it to be as
fast as their Windows and OS/2 ports, which at that time was the
fastest Just In Time compilers on the PC-platform.  It performs very
well on Linux, and the installed version is just a maintainance
upgrade.

MySQL and the corresponding JDBC driver \textit{must} be the latest
stable versions due to security and bug fixes.


\subsection{Aalborg}
\label{sec:aalborg}


(4-CPU UltraSparc 440MHz with Sun JDK 1.2)



%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "rapport"
%%% End: 
