
\chapter{Overview of technologies available for Linux February 2000}

% \framepage{15cm}{
% Describe what products there are currently available for these
% categories. Evaluate as much as possible.   Conclude that currently
% this is an area in great development.
% }

This appendix lists some of the technologies I have been looking at
while working with Cactus and possible filters with it.   This is by
no means complete.


% \section{Webservers}
% \framepage{15cm}{
% Apache, misc Servlet Javaservers (jetty, thttpd, java webserver,
% etc).  Look for comparisons.  Roxen (graphs).  Discuss whether several
% webservers should be run on the same machine to provide better services.

% Look at the netcraft search - differentiate between Internet server
% and intranet server, and internal product stuff.
% }

%\section{Search engines}



\section{Database engines}

% \framepage{15cm}{
% Full SQL servers:  MySQL/mSql/DB2/Oracle (spank their FLUG
% policy)/Informix/PostgresSQL (discuss features:  transactions, unique
% numbers, in-kernel code, speed, availablity, multi-host
% capabilities?). Any flat file RDBS?

% Tricks: Menu system in LDAP (lr.dk).
% }

The number of database vendors which support Linux (which as for now
is \textit{the} Open Source operating system which people know about)
is steadily increasing.  \myurl{http://linas.org/linux/db.html}{This
  webpage try to keep track of them all}, but I only list the major
commercial vendors and Open Source projects, where the drivers should
be of production quality.  The following overview was created March
2000.

% \framepage{15cm}{
%   PostgresSQL (discuss features:  transactions, unique
%   numbers, in-kernel code, speed, availablity, multi-host
%   capabilities?). Any flat file RDBS?

%   Greenspuns evaluation of embedded webservers in databases
%   }

\subsection{Cloudscape - Informix}
\label{sec:cloudscape}

The
\myurl{http://www.informix.com/informix/press/1999/dec99/cloudscape.htm}{Cloudscape
  Database} is written in Java, with a JDBC and a HTML interface, has
many core SQL and Java extensions implemented, and allow distribution
amongst several Java-machines (using RMI).  Informix bought Cloudscape
Inc. in October 1999 to get this database system, so it is still very
new.  They offer
\myurl{http://www.cloudscape.com/Evaluations/index.html}{a 60 day
  trial version from their website}.  Pricing is approximately \$900.


\subsection{DB2 - IBM}
\label{sec:db2}

\myurl{http://www-4.ibm.com/software/data/db2/linux/}{DB2 for Linux}
is available - the \myurl{DB2 Personal Developer's Edition
  V6.1}{http://www6.software.ibm.com/dl/db2pde/db2pde-p} is available
for free for non-commercial purposes.  

\myurl{http://www-4.ibm.com/software/data/db2/extenders/xmlext/}{DB2
  XML Extender} is an interesting addition to DB2 which ``let you
store XML documents in DB2 databases and new functions that assist you
in working with these structured documents.''.  It is available for
AIX, Solaris and Windows NT. 

  

% \subsection{Informix - Informix}
% \label{sec:informix}

% Informix is also a 
% http://www.informix.com/datablades/dbmodule/informix1.htm

 
\subsection{Ingres - Ingres}
\label{sec:ingres}

The
\myurl{http://www.cai.com/products/betas/ingres_linux/ingres_linux.htm}{Ingress
  II database is in a beta stage for Linux}, and can be downloaded for
free.  It uses Perl, gcc and Apache to provide webserver facilities.


% \subsection{mSQL - ?}
% \label{sec:msql}

% \textsf{Yes?  Looook for stuff - rememebr to get the mSQL/MySQL book
%   in the databse}

\subsection{MySQL - TCX}
\label{sec:mysql}

\myurl{http://www.tcx.se}{MySQL is an Open Source database} which is
very popular amongst Perl programmers due to
\myurl{http://www.tcx.se/benchmark.html}{the good performance}
combined with the free availability of source code and high quality
drivers.

MySQL only costs money if you run it on a Microsoft operating system,
or commercially on a server.

\subsection{Oracle}
\label{sec:oracle}

The complete Oracle 8i product range is available for Linux, with a
low-end version available for download.  The WebDB tool functions as a
webserver-interface to any Oracle database.

When
\myurl{http://www.flug.dk/pages/foredrag/tidligere_foredrag/tidligere_foredrag.html}{Oracle
  visited FLUG} we were promised that we
could have all the Oracle software for Linux we needed (without
official licenses), and that Oracle Denmark would provide the
necessary support.  It is interesting that Oracle Denmark unofficially
seed the Linux communities with unlicensed copies of their software.

\subsection{SyBase - Sybase}

Sybase \myurl{http://www.sybase.com/products/linux/}{has Sybase
  Adaptive Server Enterprise and SQL Anywhere Studio} available for
  Linux for free as long as they are used for development.  For
  production installations a license must be bought.  An earlier
  version is unsupported but may be used without restrictions for
  production enviroments.
  
  Their web integration tool does not appear to have been ported to
  Linux yet.  The \myurl{http://linas.org/linux/db.html}{Linux
    database page} lists several third-party tools which provide web
  functionality.

  Even though Microsoft does not provide Linux drivers for the
  Microsoft SQL-server, the freely available SyBase drivers for Linux
  appear to be usable.

\subsection{Other approaches to database storage}

\myurl{http://www.lr.dk}{Landbrugets r{\aa}dgivningscenter} uses a
\myurl{http://www3.innosoft.com/ldapworld/ldapfaq.html}{Lightweight
  Directory Access Protocol} serverto provide easy access to commonly
used web elements.


\section{Available browsers}
% \framepage{15cm}{
% MSIE5 (XML support, Solaris), Mozilla (alpha), Netscape (too old),
% HotJava 3.0 (no XML, can be expanded to do XML conversion
% internally?), Opera (?), Amaya (other W3C browsers?), AWTviewer in fop
% (can it do xml directly?),
%}
%\textsf{Consider rewriting this chapter!}

\subsection{Netscape Communicator}
\label{sec:netscape-communicator}

The Netscape Communicator browser is available and fully supported for
Linux.  The version available at March 2000 was 4.72 in English only.

This browser have a reputation of being rather picky of its
environment.  Especially the Java-subsystem is prone to crashing the
server.  If Java is important, consider HotJava.


\subsection{Mozilla - Netscape 6.0}
\label{sec:mozilla}

\myurl{http://www.mozilla.org}{The Mozilla browser} has been under
development since 1998, where Netscape -- inspired by the \myurl{http://www.tuxedo.org/~esr/writings/cathedral-bazaar/}{the
Cathedral and the Bazaar} paper -- decided to switch development
strategy for its incomplete source for Netscape Communicator 5.0 from
an in-house system to a full-scale OpenSource model housed at
\myurl{http://www.mozilla.org}{http://www.mozilla.org}.

In the past two years, the OpenSource developers have basically
rewritten the Communicator from scratch into a truly impressive
product (this view is based on the M14 milestone release).  A seperate
beta-branch of the source code designated for the Netscape release was
opened March 2000. A tentative release is midsummer 2000.

\subsection{Opera}
\label{sec:opera}

This \myurl{http://www.opera.com}{Norwegian browser} has been highly
successful on Windows for providing a small and quick browser which
goes well with users who think that Netscape and Internet Explorer are
too big and slow for their machines.  Their Linux port is still in
Alpha-testing, and unsuited for mainstream use.


\subsection{HotJava 3.0}
\label{sec:hotjava}

HotJava is a Netscape 3.0 compatible browser written in Java.

It is a bit slow, especially in screen updates, but has its force in
its ability to execute Java applets (which is where Netscape is slow,
and Microsoft does not comply with the standards).

HotJava requires a Sun JDK-based Java implementation (clones do not
provide all the functions they need), and I have tested version 1.1.2,
1.1.4, and 3.0 under Linux, OS/2, Windows, Irix and Solaris, where it
worked satisfactory.

There is currently no information regarding the status of parsing XML
internally.  Full access to the source is probably necessary before
any third-party implementations come.  In a perfect world, Sun would
release the source, but that is probably very unlikely.

Rumours has it that development on this browser has ceased.  I tried
to extend version 1.1.4 to show TIFF files, but had to give up due to
the miserable documentation for HotJava combined with lack of the
source.

Other rumours has it that Sun will release the source for HotJava
under their Community License (which basically makes the developers
unpaid employees of Sun) like it has been done earlier.  On March 13,
Sun released their first product -- Forte (previously Netbeans) --
under a true OpenSource license, which could indicate that they had
evaluated their experience with previous releases, and found that
nothing short of full OpenSource will do.

Remark: Today there are so many OpenSource projects that the truly
talented coders can choose freely what they want to develop on, and
they choose projects where they can get return in recognition from
their equals, and where they don't feel that they slave for a company.

Sun has a
\myurl{http://developer.java.sun.com/developer/bugParade/top25bugs.html}{``vote
  for the bug you want fixed''} system for Java, and the highest rated
bug was ``Port Java to Linux'' for a \textit{very} long time, and with
four times the number of votes for number 2.  The
\myurl{http://www.blackdown.org}{Blackdown Team} ported Java 1.1 to
Linux, and collaborated with Sun on porting Java 1.2 to Linux.  It
took very long time, and when they finally had it officially published
on the Sun Java webpages, Sun did not give credit to the Blackdown
Team -- not on the web pages, not in the code [rumours have it they
even removed comments crediting Blackdown].  The Blackdown Team was
highly offended, said publicly so, and stopped working on the port!
[As Java3D code has been made available on the Blackdown site in
March 2000, it appears that the matters have been resolved].

My personal guess is that this incident taught Sun that they must be
very observant to the OpenSource culture they want to attract.  It
must be nurtured and credited in order to create the symbiosis that
the Mozilla project has demonstrated to be possible.

In the same spirit \myurl{http://www.sourceforge.net}{SourceForge}
are providing free hosting for OpenSource projects, and they have
reached the point where people ask ``why don't we put this on
SourceForge to get CVS''.

\subsection{Microsoft Internet Explorer 5.0}
\label{sec:microsoft-internet-explorer}

Microsoft has created a very nice browser which works well on Windows
platforms.  The MSIE 5.5 beta supports a recent version of XML with
the XSL-stylesheets but apparently is not implementing it correctly.
Since the XSLT standard was released January 2000, a release of MSIE
can be expected ``rather soon'' which implements the XSLT standard,
allowing the browser to render XML files on the client side.  For now,
server side processing still appears to be the only option.

Microsoft Internet Explorer 5.0 is \textit{not} available for Linux.
It is, however, available for Solaris and HP-UX so the codebase is
ported to Unix, and from there it is usually comparatively easy to get
software to run under Linux.  The Solaris version was incompatible
with the window manager I was running under Irix, so I have not tested
it much, but it was very, very similar to the Windows version.

My personal guess is that Microsoft holds back Linux versions of the
Internet Explorer with Outlook Express until they see a fortunate
opportunity to release it.  \myurl{http://www.idg.net/microsoft}{A
  certain lawsuit springs to mind} - I do not expect Microsoft want to
be accused of monopolizing the Linux market too.

To summarize:  Microsoft Internet Explorer is not available for Linux
at the time of writing.

\subsection{Amaya}
\label{sec:amaya}

Amaya is the W3C testbed for HTML-development.  Is a nice browser
combined with a reasonable editor.  Version 2.4 do XHTML and HTML-4.0
with CSS but not XML yet.

% \subsection{AWTviewer in fop}
% \label{sec:awtviewer}

% \textsf{probably for fop stuff only}


\subsection{StarOffice}
\label{sec:staroffice}

StarOffice include a Netscape compatible browser, with
Java applet support.  It is an integrated part of the StarOffice
program, which besides Linux is available for Windows, OS/2 and
Solaris.

As Sun recently bought StarDivision who had written StarOffice, the
software currently being retargetet to Sun interests.  It is not
evident yet what that might be, but it is not unreasonable that the
thin Java client (connecting to a StarOffice server on a Solaris
machine) is what Sun bought it for.

Currently there is no sign for integrating XML support in StarOffice.

\subsection{W3 - browsing inside Emacs}
\label{sec:w3-browsing-inside-emacs}

The \myurl{ftp://ftp.cs.indiana.edu/pub/elisp/w3/}{W3-browser for
  Emacs} (needs XEmacs to show images) is a reasonably good browser,
which has its force in its tight integration with Emacs, and is an
interesting alternative for those who live in Emacs anyway.  W3 is
written in Lisp.  The next major version is expected to be rewritten
around a generic XML parsing engine.


\subsection{Lynx - browsing in textmode}
\label{sec:lynx-browsing-in-textmode}

Lynx is one of the oldest browsers around still in widespread use, and
does text-mode only.  It is being steadily improved and hacked on,
both because it is a nice browser for those who may not have a
graphical login available, but also because it has a lot of options to
help the programmer debug webservers and scripts.  Has only recently
been superseded in usefulness by the various tools in the Perl
community.


% \section{XML utilities}
% \framepage{15cm}{
% XT, Xerces, Cocoon, SQL2xml, validators (error recovering?),
% converters to XML (pod2docbook, tex4h, Perl with appropriate modules?,
% wordview modified)?

% Show the XML->SQL->XML->HTML on the fly conversion possible with
% Cocoon .... Without auxillary code. All done in W3C style.

% WordPerfect 9 should be able to edit SGML directly.
% }





% \section{PDF utilities}
% \framepage{15cm}{

% PDF generated by pdflatex is good.  PDF generated by fop isn't.
% createpdf.adobe.com 

% pdfzone.  acroread (ok, heavy), xpdf (font problems with bitmaps, ugly
% but fast, gives best result with usepackage{times}, no anti-alias), gv
% (pretty good).  Distiller (check what it can do).  Use pdfinfo to get
% the number of pages in the pdffile.

% Conclude that it is possible to have a 100\% Java solution for
% on-the-fly publishing XML to HTML and PDF through a webserver.
% Discuss why developers do it in Java, as opposed to any other language.

% Very few utilites support creating PDF.  Discuss how to create PDF.

% }

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "rapport"
%%% End: 
