% $Id$

\chapter{The DocBook and TEI DTD's}

\section{DocBook considerations}
\label{sec:docbook}

Having discussed the merits of SGML and XML, the next step is finding
and deciding on a suitable DTD for the documents.  My findings are
that there currently is two trends on the Internet regarding
DTD-development:

\begin{itemize}
\item Small DTD's designed for a specific purpose
\item Large DTD's designed for a wider array of possible applications
\end{itemize}

Since we are interested in translating to and from this document
format, and we prefer not to have to maintain these translations
ourselves, it is only the large DTD's that are interesting.

I have been looking for such large, well-maintained DTD's on the
Internet for almost a year now, and this is an extra-ordinarily rare
thing.  I have basically only found three possible candidates:

\begin{description}
\item[\myurl{http://www.docbook.org}{DocBook}] -- A DTD designed for
writing technical documentation.  It is being used by a steadily
increasing number of companies and OpenSource projects.


\item[\myurl{http://www.uic.edu/orgs/tei/}{TEI}] -- Text Encoding
Initiative.  A DTD intended to be used with ``preparation and
interchange of electronic texts for scholarly research, and to satisfy
a broad range of uses by the language industries more generally.''

\item[\myurl{http://www.openebook.org/faq.htm}{eBook}] -- the eBook
  format is intended for ``Authors, editors, publishers and content
  owners who want to have their titles in a format that is "eBook
  ready," which then can be used by a variety of electronic book
  publishing systems and reading devices.''
\end{description}

% \textsf{
% The SGML is converted to HTML by applying a HTML style sheet for
% DocBook, and sending it through jade with request for a sgml
% conversion.  The stylesheet generates a HTML file pr section in a
% chapter which is rather too much.  The HTML is \textit{very} ugly - all
% induced line breaks are \textit{within} the tags to avoid introducing any
% artificial whitespace.  The HTML is Lynx-compatible.


I have concentrated on the DocBook format simply because it had the
best documentation (DocBook: The Definitive Guide which was published
in late 1999 \cite{walsh-muellner:docbook-the-definitive-guide}), and
appeared to fulfill the needs for the document conversion utilities I
wanted to have.

Even though DocBook originally was designed for SGML, the principal
developer Normal Walsh, has done a lot of work on developing two XML
versions -- one with full functionality, and the other with a limited
subset intended for rendering inside a XML-capable browser.   This is
intended to be fully present when the official version 4.0 version of
DocBook is released around third quarter 2000.




% nice in Netscape.  A ``up'', ``next'' and ``previous'' are available
% combined with a ``Top title''-$>$''Chapter title'' -$>$ ``Section title''
% navigation bar at the top.

% The localization code is manually maintained.

It is possible to have a DocBook XML-file which can be processed both
with XSL-stylesheets \textit{and} DSSSL-stylesheets, by letting the
!DOCTYPE header point to the DocBook XML DTD along with
system-dependent path (which is required by XML).  A sample header for
DocBook is listed below:

\begin{verbatim}
<?xml version="1.0" encoding="ISO-8859-1" standalone="no"?>
<!DOCTYPE article 
  PUBLIC "-//Norman Walsh//DTD Simplified DocBk XML V3.1.3.6//EN" 
  "/home/ravn/sgml/docbk/db315/docbookx.dtd">
\end{verbatim}

If the \texttt{standalone="yes"} field is set in the \tag{?xml?}-tag
instead, it means that the XML-tools will not validate the contents
against the DTD, allowing both for faster performance, but also
without requiring anything installed on the rendering machine but the
style sheets.  Since a full installation of DocBook with all
supporting DTD's and entity-lists is a large and complex affair, this
is a very pleasant feature.

\begin{table}[htbp]
  \begin{center}
    \begin{tabular}[tb]{|l|p{12cm}|}
\hline\hline
Format & Conversion command from original (\$\(<\)) to destination (\$@)\\
\hline
  XHTML-1.0 & java com.jclark.xsl.sax.Driver \$\(<\) ../docbook/xhtml/docbook.xsl \$@\\
  HTML-4.0 & java com.jclark.xsl.sax.Driver \$\(<\)
  ../docbook/html/docbook.xsl \$@ \\

  XSL-FO & java com.jclark.xsl.sax.Driver \$\(<\) ../docbook/fo/docbook.xsl \$@ \\
  RTF & jade -t rtf -d .../docbook/print/docbook.dsl .../jade/pubtext/xml.dcl \$\(<\)\\
  MIF (Frame) & jade -t mif ... \\
  {\TeX} & jade -t tex ... Note
  that the special {\TeX}-dialect \texttt{jadetex} must be used to
  process this output.\\
  
  PDF (from FO) & \texttt{tex '\&pdffotex' document.fo}  \\
  & (Note:  This
  makes much prettier output than FOP)\\
  
  PDF (from FO) & \texttt{java org.apache.fop.apps.CommandLine \$\(<\)
  \$@}   \\
  
\hline
    \end{tabular}
    \caption{Some of the output formats possible for a DocBook XML document}
    \label{tab:output-formats-possible-for-a-docbook-xml-document}
  \end{center}
\end{table}

Table \myvref{tab:output-formats-possible-for-a-docbook-xml-document}
lists several of the possible formats which a DocBook document can be
rendered to.  See \myvref{sec:xml-publishing} for a functioning set of
Cactus filters implementing a DocBook publishing engine.

Most of these transformations are currently rather slow, currently
making them unsuitable for ``on-the-fly'' rendering.  XSL-conversions
are feasible (those implemented in Java), but the
DSSSL-transformations are simply too slow.


% \section{TEI -- \myurl{http://www.uic.edu/orgs/tei/}{Text Encoding
%     Initiative}} --
% \label{sec:tei}



% \section{Converting documents to other formats}
% \label{sec:docbook-converting-documents-to-other-formats}


% \subsection{SGML to XML}
% \label{docbook-converting-sgml-to-xml}

% \begin{alltt}
%   sx -biso-8859-1 -xlower input.sgml $>$ output.xml
% \end{alltt}

% The DTD for the document \textit{must} be registered in the
% \textsf{catalog}.  Even so a number of warnings and errors may be
% printed, since some SGML-constructions cannot be represented in XML,
% and some entities are unknown.  This must be handled manually - SX is
% not a tool do this perfectly.


% \subsection{XML to RTF}
% \label{sec:docbook-xml-to-rtf}

% Use \texttt{jade} with the appropriate stylesheet and \textsf{....}

% \begin{alltt}
%          jade -t rtf -d docbook.dsl xml.dcl inputfile.xml
% \end{alltt}

% where \texttt{docbook.dsl} is one of the stylesheets from the
% \textsf{DocBook Modular StyleSheets}
% (\texttt{dsssl/docbook/print/docbook.dsl}) and \texttt{xml.dcl} comes
% from the Jade distribution (\texttt{pubtext/xml.dcl})

% The resulting RTF-file is compatible with Word-97, Word-95 (claimed by
% author), StarOffice, \textsf{other texteditors} 

% \textsf{other backends}

% SQL to XML :: Coocncococns sql processsor

\section{A sample DocBook XML document}
\label{sec:amanda-readme.xml}

I have marked up a DocBook XML version of the
\myurl{ftp://ftp.amanda.org/pub/amanda/README}{README file for the
  AMANDA backup system}, as part of the general transition to an
SGML-based documentation set for AMANDA.  I have used it as a test
document for the XML-publishing sample filter-set in
\myvref{sec:xml-publishing}.

There are a few caveats for HTML-knowledgable people:

\begin{itemize}
\item Each paragraph must be enclosed in a \tag{para}-tag.  Also
  inside a \tag{listitem}-tag.
\item \tag{sect1}, \tag{sect2} etc. encloses around the \textit{whole} section, and the
  \tag{title}-tag shows the title of the section.
\end{itemize}  
In order to be processable by SGML tools like Jade (to get RTF output)
it is necessary to have a \tag{!DOCTYPE} entry, which again requires
-- for XML compatibility -- that the DTD is explicitly given.  This
particular document conforms to
\myurl{http://www.nwalsh.com/docbook/simple/index.html}{the
  "Simplified" DocBk XML DTD}.



\listinginput{1}{../sample/AMANDA-README.xml}


\section{Making it as easy as possible for authors to write in DocBook}
\label{sec:making-it-as-easy-as-possible-for-authors-to-write-in-docbook}

The multitude of different programs necessary to convert documents to
and from DocBook means that the author has to use several widely
different tools in order to do the different tasks which a modern word
processor like Microsoft Word have built in as standard.  Since the
original idea behind recommending a *ML dialect for the original
format of the documents, was to allow the author to write without
worrying about how the document was to be presented, it would follow
that spirit to allow the author to publish her works with as little
work as possible.  This means automation of the various underlying
processes for converting to HTML, RTF, PDF etc.

This could be done quite easily with installing all the necessary
programs on the authors computer, and providing a few scripts to call
in order to do a given task, but still leaves the author to do these
manually depending on the output format needed, and with the job to
keep the software installation current and correct.   This means a lot
of irrelevant maintainance, which the author could put to better use
by writing instead.  

Another solution in a networked environment which also is in
compliance with the finding in
\myvref{sec:the-user-should-use-familiar-tools-to-publish-documents},
namely that publishing a document to the web, ought to be just as easy
as printing a document to paper, would be that the author has access
to a DocBook publishing engine, which accepts DocBook documents (in
either SGML or XML) and uses the very best tools to produce the best
output available.  In order to be as usable as possible for most
authors it must be very easy to use, and usable without \textit{any}
special software.  Very little software can be considered available to
everyone even today, but I do not find it unreasonable to expect that
an author using a computer has a browser installed and has access to
email.

A good solution for such an author would be a system, where the
document could be submitted by email and converted to many formats,
where the converted result could be retreived with a browser from a
web site, without the author having to worry at all about the
conversion process.  Such a system should be managed centrally and be
easily extendable and customizable, in order make it easily
maintainable.

The sample system for this thesis -- named \textit{Cactus} (see
\myvref{cha:cactus}) -- does exactly this, and is generic enough to be
usable for a lot of other purposes depending on the actual
configuration.

That alone does not make the system generic enough to make it work
well in a networked environent, but building upon the framework of
a modern SQL database provides a lot more flexibility than just
working on a flat file system.   For this reason Cactus was developed
to run on top of a MySQL relational database.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "rapport"
%%% End: 
