\maketitle

\section*{Abstract}
\label{sec:abstract}


\begin{quote}
  This thesis discusses database backed websites; their current use as
  presentation engines and discussion of several sample database
  backed websites.  The weakness of the current situation is
  discussed, as many sites today are designed for \textit{visual}
  presentations for humans, which makes them useless in settings where
  the visual representation is inappropriate for the recipients.  This
  includes blind people and indexing search engines, because important
  meta-data about the web-pages are discarded.  The expected
  popularity of the WAP-technology hasten the need for websites where
  the content is seperate from the layout, allowing the browsers to
  retrieve the meta-data and content of a web page, and render it
  according to the abilities of the recipient.
  
  Technologies to do so are discussed (with an emphasis on Open Source
  implementations), and two sample systems -- \textit{Cactus} and
  \textit{Konsensus} -- are presented, and discussed.
  
  Konsensus has been almost fully implemented by using existing
  OpenSource software to allow an open group of people to maintain a
  Frequently Asked Question document for a given subject.

  Cactus is a application framework based on a database, which can be
  used for many purposes.  The proof-of-concept implementation
  described here, is envisioned to be used in a small workgroup where
  there occasionally is a need to convert documents and images due to
  software incompatabilities with the ``outside world''.

  The future home page of the Cactus project will be
  \myurl{http://cactus.sourceforge.net}{cactus.sourceforge.net}.
\end{quote}


\section*{Acknowledgements}

Many people have provided inspiration, comfort, help, and much more
for me while doing this.  I would in particular like to mention a few:

\begin{description}
\item[Astrid] -- for being a fantastic support, for being there and
  for doing just about everything related to the real world for me
  while writing.  I couldn't have done it without her.
  
\item[Laila] -- for providing feedback on several drafts on both
  content and language, as well as
  providing support and company.

\item[Matt, Chris, Brian and Burning] -- for support and
  company during the nights.
  
\item[Anders] -- For doing the figures, along with all the rest.
  
\item[Jan] -- for providing feedback and providing the graphic for
  home banking.

\item[Michael, Thomas, Steffen and Peter] -- for bearing with me when
  I was ranting as well as when I was being carried away by the music.
   They also found themselves being listeners when things went well.  


\end{description}



%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "rapport"
%%% End: 
