
\chapter{Introduction}



%\textsf{This section is rudimentary -- it must be written at the end.
%  Currently there is just keywords.}

%\textsf{In this thesis I refer to the author Annie, and the layouter
%  Larry.}

\section{What is a ``database backed webserver''?}

A \textit{web-server} is a program which hands out files to any
web-browser that asks.  Note that the term is not used in this thesis
to designate the machine running the web server, but just the program.

Phillip Greenspun names a \textit{database} as ``a spread-sheet that
several people can update simultaneously''
\cite{phillipandalexsguidetowebpublishing}.  This is, however,
slightly inaccurate -- that is more accurately used about a single
table in the database).  A \textit{database server} is a database
which allows programs running on other computers to use it.

A \textit{database backed webserver} is a web-server which knows how
to talk to a database server while servicing bypassing web-browsers.


% In order to be useful it must be \textit{fast}!

% A \textit{web server} is a program which waits for requests from web
% browsers (like Netscape Communicator) and returns do

% A \textit{databased backed webserver} is a webserver which has access
% to a database at the same time as it serves documents to clients.
% Such access opens new possibilities, some of which are:

% \begin{itemize}
% \item The database can protect against race conditions when data is
%   updated on the server.
% \item Provide a searching facility more complex than grep.
% \item Allow fast access to data without worrying about limitations in
%   file systems
% \item ``\textsf{users can extract data on their own if they want to}''
% \item Databases are optimized for disk I/O, webservers for network
%   I/O.  Those go well together.
% \item Very large amount of data can be online
% \item Data can be managed remotely by others than the webmaster
% \item \textsf{what else?}
% \end{itemize}


\section{Why use a webserver in combination with a database?}
\label{sec:why-use-a-webserver-in-combination-with-a-database}

In one word: \textit{Synergy}!


The web browser provide a simple, efficient and well known user
interface available on most computers connected to the Internet, and
the database provide efficient dynamical access to data.

The combination of the two allow for the personalized Internet where a
given webserver can give any user customized information upon request.
Examples are:

\begin{itemize}
\item \textbf{Newspapers} -- the user may specify that she wants
  in-depth coverage of international affairs, and do not want sports,
  and her personal web-edition will then just contain that.  Payment
  may be per-article instead of per-paper.
  
\item \textbf{Maps/Airlines/Hotels} -- the user may request the best
  way to travel from A to B at a given time, see the route, and order
  accompanying reservations on-line.

\item \textbf{Real stores} -- the user may browse and order from the inventory.
  The ordered goods are then delivered by mail, delivery boy or
  similar.
  
\item \textbf{Virtual stores} -- the user may browse and order from a
  virtual inventory.  The order is then analyzed and forwarded to the
  appropriate suppliers to the virtual store.
  
\item \textbf{Banks} -- there is a growing need from customers to be
  able to do home banking.  If the bank offer a browser based
  solution, they can support customers regardless of their choice of
  computer.  The transactions done through such a system will most
  likely be stored in a database to avoid data loss.
  
\item \textbf{Program development} -- The
  \myurl{http://www.mozilla.org}{Mozilla project} has shown the
  benefit of up-to-date information regarding just about anything
  related to the source code.

%\item \textbf{Public Information} -- The Danish government is 
\end{itemize}


Since its inception in 1993 the web has grown to be the ``good enough
for most things'' graphical user interface, capable of showing text
and graphics as well as provide navigation.  Due to the simplicity of
HTML it is not hard to build a plain basic client\footnote{Plan basic
  client implies no Java, no Javascript, no cookies, no plugins, no
  cascading style sheets, no HTTP-1.1 and almost everything else that
  characterize the modern browsers from Netscape, Microsoft, Sun and
  others.  However, if a webpage is well-written it can
  still be displayed on such a plain basic client. } meaning that just
about any modern computer with Internet capabilites have at least one
browser available.



% \subsection{Protection against race conditions}
% \label{sec:protection-against-race-conditions}

% A database can provide atomic operations on its data, as opposed to a
% stock Unix file system, where data unwittingly can be corrupted if two
% processes try to read and update information at the same time.  The
% concept is well known from multi-programming \textsf{the eskimo book?}

% Companies like ``valueclick'' and ``\textsf{doubleclick?}'' sell
% banner ad's (the \textsf{480x120?} pixel advertisements allowing you
% to click through to the advertisors website if you are interested).
% These are either sold by the number of \textit{impressions} (shown to a
% user) or \textit{clicks} (where the user actually clicks on the
% banner), and these must be counted in order to document that the
% customer gets what he pays for, and the banner ad provider does not
% expose more than what the customer paid for.

% Here is a database system essential in order to protect the data from
% being accidentially overwritten.  ValueClick states that they have
% \textsf{7 million impressions}  daily, which is around 80
% impressions pr second.

% \textsf{And what else is new?}

% \subsection{Provide searching facilities}
% \label{sec:providing-searching-facilities}

% A flat file basically only allow you to search the content and the
% filename with modifications based on the timestamps of the file.
% Searchings normally happen linearly through files and directories.

% Since a database may have a lot more attributes, queries can be much
% more sophisticated, and may even be done against indexed data.

% \subsection{Allow fast access to data without using file systems}
% \label{sec:allow-fast-access-to-data}



% \subsection{????}



% \subsection{Databases are optimized for disk I/O, webservers for network}
%   I/O.  Those go well together.
% \subsection{Very large amount of data can be online}
% \subsection{Data can be managed remotely by others than the webmaster}
% \subsection{Others?}


% \framepage{15cm}{
% Provide background information: Quick overview of the technology
% [webserver, database, http, sql, dynamically generated content vs]
% with a graph.  Give simple example.  Explain \textsl{when} things
% happen (on-the-fly vs statically generated pages).  Explain the
% advantages and disadvantages databases have over flat filesystems
% [transactions(protection, atomicity), extra attributes, speed, indexed
% columns, complex queries, scalability(linear search in filesystems,
% multihost db's),]
% }

\section{The new role of the webserver}

The world wide web was originally designed to \textit{present
  information} written by scientists to scientist, and this has
generalized to the extend that a modern company must consider their
website to be an important part of the image of the company.  Many do
this by providing electronic version of their presentation material,
some use it to provide online editions of their manuals and other
printed matter, as well as downloadable editions of software.  Very
few websites provide facilities that allow them to be used as a
\textit{tool} in the daily lifes of the users, similar to their word
processor, calculator and spread sheets.

The Cactus system described in \myvref{cha:cactus} is meant to fill
that gap, by providing a web based service which -- hopefully -- allow
users to exchange information regardless of the software they use.

% \framepage{15cm}{
% Compare original functionality with static HTML-files with content and
% a few CGI-scripts, to the current dynamically generated sites with
% many hits pr day.  Since HTML is generated and provides little
% abstraction it is hard to use on a higher abstraction level, and with
% unreadable code in ASP and PHP3 it is harder.  CSS was not the answer
% since it did not provide any abstraction from the presentation
% (seperate content from layout).

% Different needs of the user - wap/pdf/html/braille.  CPU, storage,
% webmaster time prevents all formats being pregenerated.
% }

% Documents must be written in a format describing \textit{content} and
% not layout.



\section{How to read this document}
\label{sec:how-to-read-this-document}

This document contain a lot of abbreviations, specialized terms,
references to web sites, figures and tables.  The abbreviations and
terms are defined in \myvref{cha:terms-and-concepts}, figures and
tables are hyperlinked from the text to their actual location, and
external references are associated to the text via a footnote, which
lists the URL.


To get the most benefit of the hyperlinks, I recommend having an
Acrobat Reader open on this document while reading a paper version.
This document is available on the web at
\myurl{http://unixsnedkeren.dk/ravn/speciale/rapport.pdf}{unixsnedkeren.dk/ravn/speciale/rapport.pdf}
which has been tested with Acrobat Reader 4.05 for Linux and
Solaris.  In case of problems, please confirm first that the version
of Acrobat Reader used is at least version 4.05, and upgrade as
appropriate. 

% A quick reference web page just containing the hyperlinks in this
% document, will be made available at
% \myurl{http://unixsnedkeren.dk/ravn/speciale/hyperlinks.html}{unixsnedkeren.dk/ravn/speciale/hyperlinks.html}



%%% Local Variables:
%%% mode: latex
%%% TeX-master: "rapport"
%%% End:
